---
title: "live_week1_004"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE)
```

## Creating R chunks

- shortcut: creating new chunk, cmd+option+I (MAC), ctrl+alt+I (Wind), 
Running the code: cmd+enter (MAC), ctrl+enter (Wind)

```{r}
1+5 # ctrl + enter, cmd+ enter
1+2
3-2+4
```

## Objects

- Variables:

defined: `=`, `<-`

```{r}
x = 5
vector = 1:5
name = "Sahifa"
```

```{r}
X <- 4 
x1 = 5
1a = 2
x_ = 3
c = 3
```

Printing the variable

```{r}
print(x) # using print function
x # just running the name

vector

1:100
```


## Mathematical Operations:

```{r}
1+5
1/5
1**5
5%%2
sqrt(2)
2^(1/2)
2^1/2
log(2)
exp(2)
```


## Logical Operations:

```{r}
1 == 1 # equality
1=1 # this creates errors because it considers the left hand side 1 as a name for a variable
1 != 2

3 > 5
3 < 5

is.na(x) # this is a function to check whether the variable has a missing value or NA

!is.na(x)
```

`and` ('&') or `or` ('|')

```{r}
TRUE & TRUE

TRUE | TRUE
```

### Types of variables:

- Numeric values:

```{r}
x = 4
class(x)
class(vector)
x + 5
x == 4
```

- Character values:

```{r}
x = 4 
X = "5"
name = "Sahifa" 
class(X)
class(name)
Name = "Sahifa "
name == Name

"Hello World"

X + 2

eq = "2+2"
```

- Logical values:

TRUE, FALSE, T, F, NA - missing values

```{r}
TRUE
T

T + F

T == TRUE

NA + 1
NA * 0
NA == NA
missing_value = NA
is.na(NA)
is.na(missing_value)
sum(c(1, 2, 3, NA), na.rm = T)
```


```{r}
print(x)
x
missing_value
```

# Vectors:

- set of values.
- vectors are also of three types: numeric, character, or logical
- c() (c stands for combine), seq(), :


```{r}
c(1, 4, 10 , 2.1)

c("Sahifa", "Miranda")

c(T, F)

class(c(1, 4, 10 , 2.1))
class(c("Sahifa", "Miranda"))
class(c(T, F))
```


```{r}
mix = c(1, "two", 3, "four")

c(1, "Tower", 3, "flour")

c(T, F, "TWo ", "Create")

c(1, 2, 4, T, F)

c(1, "Tower", 3, "flour", NA)

class(c(1, "Tower", 3, "flour"))
class(c(1, "Tower", 3, "flour", NA))
```

- seq() (stands fro sequence) or `:`

seq(from = , to = , by = , length.out = )


```{r}
seq(1, 100, by = 3)

seq(1, 100, length.out = 6)

1:100
pi

pi:20
```

- rep(): repeats the first argumnet second arg no. of times

```{r}
rep(5, 10)
rep(c(1, 2, 3), 10)
```


```{r}
class(c(1, 2, 4, 5))
class(as.integer(c(1, 2, 4, 5)))
```

```{r}
#as.factor()
```

# Operations with vectors


```{r}
1:10
10:20
1:10 + 10:20

1:3
4:6
1:3+4:6

c("one", "two") + c("three", "four")

c("one", "four") == c("three", "four")

(1:100)^2%%5

(1:100)^2%%5 %in% c(0, 2)

c(0, 2) %in% (1:100)^2%%5
```

```{r}
sum(1:100)
sqrt(1:100)


```



# Friday

```{r}
# run this line of code and interpret what the output is and why it is different than what we expect from the in operator

c(0, 2) %in% ((1:100)^2%%5)
```


## Dataframes 

- `tibble()`, `data.frame()`, `data_frame()`

```{r}
# dataset for all the alphabets, first column will be the index of the alphabets, second column will contain the letters

alphabets = tibble(index_alp = 1:26, alphabet = LETTERS, vowels = alphabet %in% c("A", "E", "I", "O", "U"))

alphabets
print(alphabets, n = Inf)
```

```{r}
alphabets[5, 2]
fifth_row = alphabets[5,]
alphabets[,2]

second_col = alphabets$alphabet
```


```{r}
second_col[5]
```


## Functions:

- function_name(arg1 = value, arg2 = value, ....)

seq(from = , to = , by = , length.out = )

```{r}
?tibble
```

```{r}
which(alphabets$vowels==TRUE)
```


```{r}
getwd()
setwd("/Users/s.sahifa98/Desktop/STAT240/lecture notes")
getwd()
```



