---
title: "live_week3"
author: "Sahifa"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE)
# this is a must have
library(tidyverse)
```


# Dataset

- in-built dataset called `mpg` in `ggplot2`

```{r}
# dataset

?mpg # more information about the dataset

view(mpg) # is to open the dataset in a different tab in the source

mendota = read_csv("../data/lake-mendota-winters-2025.csv")
```

# ggplot

# Step 1: building the base layer or empty canvas

```{r}
ggplot()
```

# Step 2: specifying the dataset (mpg)

```{r}
ggplot(data = mpg)

ggplot(mpg)
```

# Step 3: specifying the axes (cty , hwy mileages)

```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy))

ggplot(mpg, aes(cty, hwy))
```

# Step 4: specifying the layers (point plot, line plot, or etc)

```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_point() +
  geom_line()
```


- customization

```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_point(color = "red") +
  geom_line()

ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_point(col = "red") +
  geom_line()
```

```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_line() +
  geom_point(col = "red") 
```
# Wednesday

```{r}
cty = mpg$cty
hwy = mpg$hwy

ggplot(mapping = aes(x = cty, y = hwy, col = cyl)) +
  geom_point()
```
```{r}
ggplot(mpg, aes(cty, hwy, col = as.factor(cyl))) +
  geom_point() +
  geom_line()
```
# variable vs constant aesthetic

```{r}
ggplot(mpg, aes(cty, hwy, col = as.factor(cyl))) +
  geom_point() +
  geom_line()

ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_line() +
  geom_point(col = "red")
```

# global vs local aesthetic

```{r}
ggplot(mpg, aes(cty, hwy, col = as.factor(cyl))) + # global aesthetic
  geom_point() +
  geom_line()

ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_line() +
  geom_point(col = "red") # local aesthetic
```
```{r}
ggplot(mpg, aes(cty, hwy, col = as.factor(cyl))) + # global aesthetic
  geom_point() +
  geom_line()

ggplot() + 
  geom_point(data = mpg, mapping = aes(cty, hwy, col = as.factor(cyl))) + # local aesthetic
  geom_line()
```

# Two Variable plots:

- geom_point():
- geom_line():
- geom_smooth(): this makes a trendline or best fit line

```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_point() +
  geom_smooth()

ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_point() +
  geom_smooth(method = "lm", se = F, col = "red") # lm = linear model, se = standard error
```
- geom_col(): this is for creating bar graphs, x axis has to have categorical variables

```{r}
summary_mpg = count(mpg, class) # we will learn function in dplyr

ggplot(summary_mpg, aes(x = class, y = n)) + 
  geom_col(color = "purple", fill = "yellow")
```

