---
title: "live_week1_003"
author: "Sahifa"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Monday

## Creating R chunks

- shortcut: creating new chunk, cmd+option+I (MAC), ctrl+alt+I (Wind), 
Running the code: cmd+enter (MAC), ctrl+enter (Wind)

```{r}
1+5 # shortcut of running an R code cmd+enter (MAc) ctrl+enter (windows)
```

# Mathematical operation:

```{r}
1+5
1-5
3 - 2 +4

1:3 + 2:4

2^3

sqrt(2)
log(1)
1:100
```
## Logical Operations:

```{r}
3 > 5
3 == 5 # equivalency or equal-to
1 == 1
1 != 1 # not equal
2 <= 4
2<=4
```

`and` ('&')  and `or` ('|'): the mathematical 'and' and 'or' operators

```{r}
T & T
T & F
F & T
F & F
```

```{r}
T | T
T | F
F | T 
F | F
```



# Objects

### Variables

Anything defined with an `=` or `<-` sign in R is a variables

```{r}
x = 5
```

```{r}
X <- 4
```

```{r}
x1 = 4
x_ = 3
x- = 2 # create an error
x. = 2
1x = 2
2*x # this is what the previous line of code expects
```

```{r}
x = 5
name = "Sahifa"
eq = "2+2"
vector = 1:10
```

## Types of objects

- Numeric values

```{r}
x
X = 4 
vector
class(x)
class(vector)

X == 4
```


- Characters:

Normal text saved as variables. All character variables need to be enclosed in quotations.

```{r}
Name = "Sahifa "
name == Name
a = "2"
b = "3"
a + b
```

```{r}
"Hello World"
```

# Wednesday

### Logical Variables:

TRUE, T, FALSE, F, NA

```{r}
x = 4
t = TRUE
f = FALSE
t + f
n_a = NA # missing values, stands for not available

NA == n_a
NA+1
```

# Vectors:

- Set of values saved as a whole

- Numeric, Character, and logucal vectors

- `c()`: c - combine

```{r}
c(1, 3, 6, 8, 10)
numbers = c("one", "three", "six", "eight")

class(c("one", "three", "six", "eight"))

mix_match = c(1,2, "three", "four")
class(mix_match)

c(1, 2, "names")


class(c(1, 2, T , FALSE))

c(1, 2, T , FALSE, NA)

c(1, 2, T, "FALSE")
```

- seq(), :

```{r}
1:100
pi:20
```

```{r}
seq(1, 100, by = 2)
seq(0, 100, by = 2)

seq(1, 20, length.out = 100)
```

```{r}
a = 1:10
b = 10:20
a
b
a + b

1:5
6:10

1:5 * 6:10

sqrt(1:10)
mean(1:10)
sum(1:10)
```

```{r}
c("1", "2", "3") + c("4", "5", "6")
```

```{r}
no. = 1:100 %% 3

# 1.5 is in no.

1.5 == no.


no. == 1
1.5 %in% no.

c(1:100, NA)
is.na(c(1:100, NA))
```


# Friday

```{r}
# run this line of code and interpret what the output is and why it is different than what we expect from the in operator

c(0, 2) %in% (1:100)^2%%5 # this was finding out the remainder of dividing the output of c(0, 2) %in% (1:100)^2 by 5

c(0, 2) %in% ((1:100)^2%%5)
```

## Dataframes


- tibble(), data.frame(), data_frame()

```{r}
alphabets = tibble(index_alp = 1:26, alphabets = LETTERS, vowel = LETTERS %in% c("A", "E", "I", "O", "U"))
```

```{r}
alphabets
print(alphabets, n = Inf)
```

```{r}
alphabets[5, 2]
alphabets[10, 2]
alphabets[10,]
alphabets[,2]
alphabets[27, 1] = 27

alphabets$alphabets
```


```{r}
alp = LETTERS
alp[5]
```


## Functions:

function_name(arg1 = value, arg2 = value,... )

```{r}
getwd() # gets working directory
setwd("/Users/s.sahifa98/Desktop/STAT240/lecture notes") # sets working directory to the path provided
```



